---
title: "LatentSNA Simulation Example"
output:
  html_document: default
  pdf_document: default
---
## Introduction
This document demonstrates a simulation study using the LatentSNA package to analyze brain connectivity data and behavioral outcomes. The simulation includes data generation, model fitting, and result visualization. The goal is to demonstrate how latent space models can be used to analyze the relationship between brain connectivity patterns and behavioral outcomes.

## Load prerequisite packages
```{r}
library(MASS)
library(psych)
library(coda)
library(magic)
```

## Source LatentSNA code
We first define a helper function to source all R files from the LatentSNA package directory. This function will load all the necessary functions for our analysis.

```{r}
#set seed and load the latentSNA code
set.seed(1)

sourceEntireFolder <- function(folderName, verbose=FALSE, showWarnings=TRUE) {
  files <- list.files(folderName, full.names=TRUE)
  
  # Grab only R files
  files <- files[ grepl("\\.[rR]$", files) ]
  
  if (!length(files) && showWarnings)
    warning("No R files in ", folderName)
  
  for (f in files) {
    if (verbose)
      cat("sourcing: ", f, "\n")
    ## TODO:  add caught whether error or not and return that
    try(source(f, local=FALSE, echo=FALSE), silent=!verbose)
  }
  return(invisible(NULL))
}

sourceEntireFolder("/gpfs/gibbs/pi/zhao_yize/xz634/latentSNA-main/R", verbose=FALSE, showWarnings=TRUE) 
```

## Generate Simulated Data
In this section, we generate simulated data that mimics brain connectivity patterns and behavioral outcomes. We set up:

* A sample of 1000 subjects
* 20 brain regions
* 5 significant brain regions with strong correlations (0.9)
* A single behavioral outcome
* One-dimensional latent space

First, we set up basic parameters:
```{r}
N<-1000 #N is sample size
ids=seq(1,N)

V<-20 # V is number of brain regions
P<-1 # P is number of behaviors

K<-1 # K and D are latent space
D<-1

W<-NULL# W and H are covariates
H<-NULL
```

Next, we create the covariance structure with significant brain regions:

```{r}
a_t<-matrix(0, nrow = N, ncol = 1)
S <- diag(1,V+D)
n_signa=V

id_siga=c(1,2,3,4,5) # sepcify significant brain regions

id=list()
id[[1]]=id

# Set correlation structure for significant regions
for (each in id_siga){
  
  S[each,id_siga[!id_siga %in% each]]=.9
}

S[(V+1),id_siga]=.9
S[id_siga,(V+1)]=.9

# Extract submatrices
Su = matrix(S[1:V,1:V], nrow=V, ncol=V)
Stheta = matrix(S[(V+1):(V+D),(V+1):(V+D) ], nrow=D, ncol=D)
Sutheta =matrix(S[(V+1):(V+D),1:V], nrow = D, ncol = V)

# Print Sutheta to show that brain regions 1-5 are significantly associated with behavior (correlation = 0.9)
print(Sutheta)
```
Generate latent variables using multivariate normal distribution:
```{r}
UTheta <- mvrnorm(n = N, mu=rep(0,(V+D)), Sigma=S, empirical = FALSE)
U.array=array(NA, dim = c(V,K,N))
U.array[,1,]=t(UTheta[,1:(V)])
Theta_t <- data.matrix(UTheta[,(V+1):(V+D)])
rownames(Theta_t)=ids
```

Set model parameters and generate connectivity matrices:
```{r}
beta_t=NULL
gamma_t=NULL

#connectivity variance
s2_t=1
#attribute variance
s1_t=1

Alpha_t=matrix(1, nrow = P, ncol = 1)
theoretical.str=Alpha_t
b_t=matrix(0, nrow = P, ncol = 1)

X<-list()
for(i in 1:N){
  
  errormatrix=matrix(0, nrow = V, ncol = V)
  errormatrix[upper.tri(errormatrix,diag = FALSE)]=rnorm(V*(V-1)/2, sd=sqrt(s2_t))
  errormatrix=t(errormatrix)+errormatrix
  diag(errormatrix)=rnorm(V, sd=sqrt(s2_t))
  
  X[[i]]<-as.numeric(a_t[i,])  + U.array[,,i]%*% t(U.array[,,i]) +errormatrix
  #diag(X[[i]])=NA
  
}
```

Finally, generate behavioral outcomes:
```{r}
Y= matrix(rep(1,N)) %*% t(b_t)+ Theta_t%*%t(Alpha_t)+matrix(rnorm(N*P, sd=sqrt(s1_t)),N,P)
```

## Fit LatentSNA Model and Store Results
Here we show the model fitting process for completeness. However, since this process takes several hours to run, we have pre-computed the results and saved them for analysis. The process includes:

1. Data Preparation:
   - Assigning proper names to observations in X and Y
   - Creating test and training sets (10% test, 90% training)
   - Storing the full dataset for later comparison

2. Model Specification:
   - MCMC parameters: 20,000 iterations with 15,000 burn-in
   - Random seed set to 1 for reproducibility
   - No additional covariates (W=NULL, H=NULL)

Here is the code for reference:
```{r, eval=FALSE}
names(X)=as.character(seq(1,length(X)))
rownames(Y)=as.character(seq(1,length(X)))
ids=names(X)

sampled.id=sample(ids, round(length(ids)*.1), replace = FALSE) # use 10% as test data, 90% as traning data

train.id=ids[!ids %in% sampled.id]
## store real data 
X_full=X
Y_full=Y

Y[sampled.id,]=NA

X.array=simplify2array(X)

#This model fitting process could take several hours
model1=latentSNA(X, Y,W=NULL, H=NULL,
                 seed = 1, nscan = 20000, burn = 15000, odens = 1,
                 prior=list())

res=list("model"=model1,"X_full"=X_full,"Y_full"=Y_full,"sampled.id"=sampled.id)
saveRDS(res,"/gpfs/gibbs/pi/zhao_yize/xz634/Revision/Simulation/Result/simulation.rds")
```

## Explanation of Results
The LatentSNA model uses MCMC to estimate parameters, producing several key outputs:

```{r}
result<-readRDS("/gpfs/gibbs/pi/zhao_yize/xz634/Revision/Simulation/Result/simulation.rds")
result$model$COV |> dim()
result$model$THETAPM |> dim()
```
1. COV matrix (20000 × 20): Provides covariance estimates between each brain region and behavior across 20000 MCMC iterations
2. THETA (20000 × 1000): Contains latent variable estimates for each subject across iterations. When there are no covariates and intercept is set to 0, these estimates directly correspond to predicted behavioral outcomes.

While the model produces other variables, we focus on these two key outputs as they are most crucial for evaluating brain-behavior relationships and prediction performance.

## Visualize Significant Brain Regions
Compare the estimated covariance structure with the true values:
```{r}
# Create data frame for plotting
brain_regions <- 1:20
estimated_cov <- abs(result$model$COV |> apply(2,mean))
true_cov <- as.numeric(Sutheta)

comparison_df <- data.frame(
  Region = brain_regions,
  Estimated = estimated_cov,
  True = true_cov
)

# Create comparison plot
library(ggplot2)

ggplot(comparison_df, aes(x = Region)) +
  geom_line(aes(y = Estimated, color = "Estimated"), linewidth = 1) +
  geom_line(aes(y = True, color = "True"), linewidth = 1) +
  geom_point(aes(y = Estimated, color = "Estimated"), size = 3) +
  geom_point(aes(y = True, color = "True"), size = 3) +
  scale_color_manual(values = c("Estimated" = "blue", "True" = "red")) +
  labs(title = "Comparison of Estimated vs True Brain-Behavior Associations",
       x = "Brain Region",
       y = "Absolute Correlation",
       color = "Type") +
  theme_minimal() +
  ylim(0, 1) +
  theme(
    legend.position = "top",
    plot.title = element_text(hjust = 0.5)
  )
```
The plot shows:

* Red line: True correlations (0.9 for regions 1-5, 0 for others)
* Blue line: Estimated correlations from the model
* We can see how well the model recovered the true brain-behavior associations

As shown in the plot, the estimated correlations (blue line) closely match the true correlations (red line), with high values for regions 1-5 and near-zero values for the remaining regions. This demonstrates that our latentSNA model can efficiently identify the significant brain regions that are truly associated with behavior.

## Prediction Accuracy(Comparison between LatentSNA with CPM)
We compare our LatentSNA model with the Connectome-based Predictive Modeling (CPM) approach:

```{r}
X.array=simplify2array(result$X_full)
Y_full<-result$Y_full
Behavior = matrix(rowSums(Y_full, na.rm=TRUE))
rownames(Behavior) = ids
sampled.id<-result$sampled.id
train.id<-setdiff(ids,sampled.id)

# Fit CPM model
library(NetworkToolbox)
cpm_model = cpmIV(X.array[,,train.id], Behavior[train.id,], 
                  kfolds=10, 
                  thresh = .01,
                  connections = "overall", 
                  groups = NULL,
                  method = "mean", 
                  model = "linear",
                  corr = "pearson",
                  standardize = FALSE, 
                  progBar = TRUE, 
                  plots = FALSE)

# Get CPM predictions
X.tmp = sapply(1:length(X), function(x) X[[x]][upper.tri(X[[x]], diag = FALSE)], simplify = FALSE)
X.mat.tmp = simplify2array(X.tmp)
colnames(X.mat.tmp) = ids

predictor = t(X.mat.tmp[as.logical(cpm_model$Mask[upper.tri(cpm_model$Mask, diag = FALSE)]),])
rownames(predictor) = ids

train.predictor = predictor[train.id,]
test.predictor = predictor[sampled.id,]

# Fix the indexing for train.model
train.model = lm(Behavior[train.id,] ~ train.predictor)
test.Y_cpm = test.predictor %*% matrix(train.model$coefficients[-1])

# Create comparison scatter plots
comparison_df = data.frame(
  Actual = as.numeric(Behavior[sampled.id,]),
  LatentSNA = as.numeric(result$model$Theta[as.numeric(result$sampled.id)]),
  CPM = as.numeric(test.Y_cpm + train.model$coefficients[1])  # Add intercept back
)

# Create scatter plots
library(ggplot2)
library(gridExtra)

# Plot for LatentSNA
p1 = ggplot(comparison_df, aes(x = Actual, y = LatentSNA)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", linetype = "dashed") +
  labs(title = "LatentSNA Predictions",
       x = "Actual Values",
       y = "Predicted Values") +
  theme_minimal() +
  annotate("text", x = min(comparison_df$Actual), y = max(comparison_df$LatentSNA),
           label = paste("Correlation =", round(cor(comparison_df$Actual, comparison_df$LatentSNA), 3)),
           hjust = 0, vjust = 1)

# Plot for CPM
p2 = ggplot(comparison_df, aes(x = Actual, y = CPM)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", color = "red", linetype = "dashed") +
  labs(title = "CPM Predictions",
       x = "Actual Values",
       y = "Predicted Values") +
  theme_minimal() +
  annotate("text", x = min(comparison_df$Actual), y = max(comparison_df$CPM),
           label = paste("Correlation =", round(cor(comparison_df$Actual, comparison_df$CPM), 3)),
           hjust = 0, vjust = 1)

# Arrange plots side by side
grid.arrange(p1, p2, ncol = 2)
```

Comparing the prediction performance between LatentSNA and CPM, our model achieves a higher correlation with the actual behavioral values than CPM, indicating that LatentSNA can provide more accurate predictions of behavior from brain connectivity patterns.