---
title: "LatentSNA for Brain-behavior Linking"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
  
---

<style type="text/css">
  body{
  font-size: 12pt;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Overview


The LatentSNA model is specifically designed for identifying imaging biomarkers using brain connectivity data. It offers robust tools for unbiased estimation of imaging biomarkers' influence on behavior variants, quantification of the uncertainty and evaluation of the likelihood of the estimated biomarker effects against chance, brain-behavior prediction in novel samples for both connectivity and behaviors.



## Installation

This toolkit is implemented in R. Follow these steps for setup:


1. Clone or download the repository to your local machine.
2. Open R and navigate to the directory containing the toolkit.
3. Run the main.R script.


```{r setup}
devtools::install_github("selenashuowang/latentSNA")
```



## Required Data


The toolkit is designed to analyze data consisting of brain connectivity and individual outcomes. Specifically, it requires two key files:

* `X`: a list of $V \times V$ brain connectivity data.
* `Y`: a matrix of $N \times P$ individual outcome data.

Additionally, simulated example data is available in the directory `data/X.RData` and `data/Y.RData` for demonstration purposes.


## Key parameters

* `W` a matrix of $N \times Q$ covariates for the connectivity data.
* `H` a matrix of $N \times Q1$ covariates for the attribute data.
* `nscan` number of iterations of the Markov chain (beyond burn-in)
* `burn` burn in for the Markov chain

Note that
